Opportunities for improvement in adult trauma patients admitted to the intensive care unit: A registry-based study

### 2024-02-21 manuscript.Rmd

For all injuries, quality hospital care can reduce the amount of short-
and long-term disability. Therefore, improving planning, access and
organisation of trauma care systems including pre-hospital care,
interhospital transportation, telecommunication and in-hospital care all
play a central role in reducing overall impact after trauma [@1].
Disability-adjusted life-years (DALYs) is a way to measure
post-discharge quality of life. Trauma accounts for one-tenth of DALYs,
most of the burden in low and middle income countries (LMIC) [@5, @6].
Trauma costs approximately 3% of gross domestic product (GDP) yearly
globally, as it mostly affects the working population [@7].

@@

Many initiatives for quality improvement programs have been seen in
trauma and critical care based on the paradigm of Donabedian [@8],
stating that implementing proven structures and processes of care is the
most effective way to improve outcomes. One example of a structure and
process intervention in the intensive care unit is highly-staffed ICUs
that has been associated with improved outcomes, as well as bundles to
recognize and treat common conditions such as those for treating sepsis
and ventilator-associated pneumonia [@9]. The intention of quality
improvement is objective improvement in regards to preventability. A
cornerstone in developing the trauma quality improvement programs is
therefore multidisciplinary mortality and morbidity review. The purpose
of these is to discuss preventable deaths and identifying OFIs in trauma
care regarding structure as well as clinical processes. This way,
identifying OFIs will provide guidance as to what to focus specific
efforts on. There are several other methods to improve trauma care.
However, mortality and morbidity review has been shown to be a key
method of addressing all necessary components of trauma improvement
[@10].

@@

There are many studies on preventable deaths in trauma patients treated
in the ICU both regarding the medical aspect as well as the human
aspect. According to a study from Japan on the medical aspect, the
primary medical causes of death after trauma are haemorrhage,
sepsis/multiorgan failure and central nervous system injury. Delayed
hemostatic procedures and transfusions were common areas for improvement
in the hospital stage [@13]. The most prominent OFI in regards to
haemorrhage was decision-making compared to errors in technical skill.
The OFIs frequently involved the decision between radiology, surgery and
further investigation [@14].

@@

@@

Many patients with severe trauma are admitted to the intensive care unit
(ICU), but little is known about opportunities for improvement in this
group of patients. This being due to the patients admitted to the ICU
being more complex. Identifying OFIs in this group can therefore be more
of a challenge. There are existing studies on different aspects of
trauma care improvement, but studies focusing mainly on OFI in the ICU
are few. The aim of this study is therefore to characterise OFI in adult
trauma patients admitted to the ICU and assess how patient factors are
associated with OFI in these patients.

@@

This paragraph is not finished! Will finish after I start working on the
real data. Will maybe add a table.

@@

Table 1: deskriptiv tabell, stratifierad från utfallet, förslag är att
använda:
<https://www.danieldsjoberg.com/gtsummary/articles/tbl_summary.html>

Table 2: huvudresultat, förslag att använda:
<https://www.danieldsjoberg.com/gtsummary/reference/tbl_regression.html>
presentera ojusterad (univariable logistisk regression) och justerad
(multivariable log reg)

Tillse korrekt presentation av resultaten i textformat:
<https://www.danieldsjoberg.com/gtsummary/articles/inline_text.html?q=tbl%20_%20re#inline_text_tbl_regression>

Ytterligare tabeller, audit filtren? En flow chart för
inklusion/exklusion?

```         
```

>>>>

For all injuries, quality hospital care can reduce the amount of short-
and long-term disability. Therefore, improving planning, access and
organisation of trauma care systems including pre-hospital care,
interhospital transportation, telecommunication and in-hospital care all
play a central role in reducing overall impact after trauma [@1].
Disability-adjusted life-years (DALYs) is a way to measure
post-discharge quality of life. Trauma accounts for one-tenth of DALYs,
most of the burden in low- and middle-income countries (LMIC) [@5, @6].
Trauma costs approximately 3% of gross domestic product (GDP) yearly
globally, as it mostly affects the working population [@7].

@@

Many initiatives for quality improvement programs have been seen in
trauma and critical care based on the paradigm of Donabedian [@8],
stating that implementing proven structures and processes of care is the
most effective way to improve outcomes. One example of a structure and
process intervention in the intensive care unit is highly-staffed ICUs
that has been associated with improved outcomes, as well as bundles to
recognize and treat common conditions such as those for treating sepsis
and ventilator-associated pneumonia [@9]. The intention of quality
improvement is objective improvement in regard to preventability. A
cornerstone in developing the trauma quality improvement programs is
therefore multidisciplinary mortality and morbidity review. The purpose
of these is to discuss preventable deaths and identifying OFIs in trauma
care regarding structure as well as clinical processes. This way,
identifying OFIs will provide guidance as to what to focus specific
efforts on. There are several other methods to improve trauma care.
However, mortality and morbidity review has been shown to be a key
method of addressing all necessary components of trauma improvement
[@10].

@@

There are many studies on preventable deaths in trauma patients treated
in the ICU both regarding the medical aspect as well as the human
aspect. According to a study from Japan on the medical aspect, the
primary medical causes of death after trauma are haemorrhage,
sepsis/multiorgan failure and central nervous system injury. Delayed
hemostatic procedures and transfusions were common areas for improvement
in the hospital stage [@13]. The most prominent OFI in regard to
haemorrhage was decision-making compared to errors in technical skill.
The OFIs frequently involved the decision between radiology, surgery and
further investigation [@14].

@@

Many patients with severe trauma are admitted to the intensive care unit
(ICU), but little is known about opportunities for improvement in this
group of patients. This being due to the patients admitted to the ICU
being more complex. Identifying OFIs in this group can therefore be more
of a challenge. There are existing studies on different aspects of
trauma care improvement, but studies focusing mainly on OFI in the ICU
are few.

@@

The aim of this study is to characterise OFI in adult trauma patients
admitted to the ICU and assess how patient factors are associated with
OFI in these patients.

@@

@@

[n] patients were included in the trauma registry and trauma care
quality database between 2017 and 2022. [n] patients were excluded,
leaving a total of [n] patients eligible for the study. [n] were
excluded because they were under 15 years old. [n] patients were
excluded because they were dead on arrival. [n] were excluded because
they were not treated in the ICU, and [n] patients were excluded because
they had missing data (?).

[n{%}] unique patients of the [n] patients included were flagged with
OFI. Table 1 illustrates the number of patients flagged per audit filter
used by the Karolinska University Hospital in Solna.

Table 2 present sample characteristics. [n{%}] patients were male and
[n{%}] patients were female. [n{%}] patients had no intubation. Most
patients of the intubated ones were intubated for xx days [n{%}]. The
number of patients who arrived outside working hours - on Saturday or
Sunday or after 5 pm or before 8 am were [n{%}]. [n{%}] patients had an
ASA-score of 1-2 before arrival to the hospital, and [n{%}] patients had
an ASA-score of 3-7 before arrival to the hospital. [n{%}] patients died
within 30 days after hospitalization.

Table 3 shows the unadjusted and adjusted associations of the selected
patient and process factors in relation to OFI. In the unadjusted
analysis the factors that were significantly associated with the
presence of OFI were: ... Male
`r inline_text(table3a, variable = Sex, level = "Male")` Female
`r inline_text(table3a, variable = Sex, level = "Female")` Age
`r inline_text(table3a, variable = Age)` No intubation
`r inline_text(table3a, variable = Intubation, level = "No intubation")`
Respiratory rate `r inline_text(table3a, variable = RespiratoryRate)`
Systolic blood pressure
`r inline_text(table3a, variable = SystolicBloodPressure)` Glasgow Coma
Scale `r inline_text(table3a, variable = GlasgowComaScale)` Jour
`r inline_text(table3a, variable = Jour, level = "Yes")` Time to first
CT `r inline_text(table3a, variable = TimeFCT)` Time to first
intervention `r inline_text(table3a, variable = TimeFInt)` ASA preinjury
1-2 `r inline_text(table3a, variable = ASApreinjury, level = "ASA 1-2")`
Death within 30 days
`r inline_text(table3a, variable = Survival, level = "Dead")`

In the adjusted analysis the following factors demonstrated significant
association with OFI: ...

x was significantly associated with increased odds of OFI in both the
unadjusted and adjusted analyses.

x were not significantly associated with OFI in the unadjusted or the
adjusted analyses.
